Space is the Place: Effects of Continuous Spatial Structure on Analysis of Population Genetic Data

### 2019-11-20 ms/review-responses.tex

    We definately think that a comparison to the spatial Lambda-Fleming-Viot
    would be an extremely useful thing to do, for the reasons the reviewer mentions,
    but this would constitute an entirely different study.
    The SLFV model is only known to be an approximation of one particular biologically explicit model
    (of patchy extinction-recolonization)
    and is only conjectured to be a good approximation
    for spatial models more generally.
    To reiterate, the definition of the SLFV was explicitly motivated by a desire to approximate
    spatial models, especially Bolker-Pacala models like ours.
    Therefore, our study is an important prerequisite
    for simulation study of the spatial Lambda-Fleming-Viot,
    as it works through some of the issues in simulating biologically realistic spatial models,
    and describes some of the patterns.

\begin{point}{\autoref{fig:genparams}:}
    I do not understand why the neighborhood size varies to the same extent in the random mating model as it does for the spatial model. For the random mating model, I would have expected the neighborhood size to be equal to the census size since all individuals have the same probability of being a parent of any given offspring. From [the paragraph at \revref], it is clear that the spatial model would converge to the random mating model when the mean parent-offspring distance tends to infinity only if we were to ignore the impact of range edges. I am thus wondering whether the variation of neighborhood size one observes in Fig 2 for the random mating model is just a consequence of border effects. If that is the case, then the authors should state it clearly and try to justify it from a biological perspective.
\end{point}

\reply{
    This is indeed part of what's going on.
    We have a added some discussion of the census size scaling \llname{censusscaling}.
}

\begin{point}{\revref}
    ``Many more species occur in a middle range of neighborhood sizes between 100 and 1000 - a range in which spatial processes play a minor role in our analyses [...]''. Do the authors think that the spatial processes would still play a minor role when neighborhood sizes exceed 100-1000 if the habitat was larger than that taken in the present simulations? It would also probably be useful to mention that neighborhood sizes given in Table 1 should be compared with extreme caution since the size of the corresponding habitats vary across species. More generally, I suspect that the size of the habitat has a substantial impact on the vast majority of statistics examined in this study. Indeed, the mean parent-offspring distance, which is at the core of the definition of Wright's neighborhood size, is only small or large relative to the size of the habitat.
\end{point}

### 2019-11-20 ms/spaceness.tex
.

To isolate spatial effects from other components of the model such as overlapping generations, increased variance in reproductive success, and density-dependent fitness, 
we also implemented simulations identical to those above except that mates are selected uniformly at random from the population,
and offspring disperse to a uniform random location on the landscape. 
We refer to this model as the ``random mating'' model, in contrast to the first, ``spatial'' model.

@@

\subsection{Where are natural populations on this spectrum?}
For how much of the tree of life do spatial patterns circumscribe genomic variation? In Table \ref{table:NStable} we gathered estimates of neighborhood size from a range of organisms to get an idea of how likely dispersal is to play an important role in patterns of variation. These values should be compared to our simulation results with some caution, as though we expect neighborhood size to have similar effects across species of varying global $N_e$ \citep{Wright1946}, in our study we evaluated only a relatively small population of $\approx 10,000$. \llabel{popsizecaveat}
In addition, these empirical examples are likely biased towards small-neighborhood species (because few studies have quantified neighborhood size in species with very high dispersal or population density). However, from the available data we find that neighborhood sizes in the range we simulated are fairly common across a range of taxa. At the extreme low end of empirical neighborhood size estimates we see some flowering plants, large mammals, and colonial insects like ants. Species such as this have neighborhood size estimates small enough that spatial processes are likely to strongly influence inference. These include some human populations such as the Gainj- and Kalam-speaking people of Papua New Guinea, in which the estimated neighborhood sizes in \citep{Rousset1997} range from 40 to 410 depending on the method of estimation. Many more species occur in a middle range of neighborhood sizes between 100 and 1000 -- a range in which spatial processes play a minor role in our analyses under random spatial sampling but are important when sampling of individuals in space is clustered. Surprisingly, even some flying insects with huge census population sizes fall in this group, including fruit flies (\textit{D.~melanogaster}) and mosquitoes (\textit{A. aegypti}). Last, many species likely have neighborhood sizes much larger than we simulated, including modern humans in northeastern Europe \citep{Ringbauer2017}. For these species demographic inference and summary statistics are likely to reflect minimal bias from spatial effects as long as dispersal is truly continuous across the landscape. While that is so we caution that association studies in which the effects of population structure are confounded with spatial variation in the environment are still sensitive to dispersal even at these large neighborhood sizes.

>>>>

To isolate spatial effects from other components of the model such as overlapping generations, increased variance in reproductive success, and density-dependent fitness,  \revpoint{3}{6}
we also implemented simulations identical to those above except that mates are selected uniformly at random from the population,
and offspring disperse to a uniform random location on the landscape. 
We refer to this model as the ``random mating'' model, in contrast to the first, ``spatial'' model.

@@

\subsection{Where are natural populations on this spectrum?}
For how much of the tree of life do spatial patterns circumscribe genomic variation? In Table \ref{table:NStable} we gathered estimates of neighborhood size from a range of organisms to get an idea of how likely dispersal is to play an important role in patterns of variation. These values should be compared to our simulation results with some caution, as though we expect neighborhood size to have similar effects across species of varying global $N_e$ \citep{Wright1946}, in our study we evaluated only a relatively small population of $\approx 10,000$. \llabel{popsizecaveat}
In addition, these empirical examples are likely biased towards small-neighborhood species (because few studies have quantified neighborhood size in species with very high dispersal or population density). However, from the available data we find that neighborhood sizes in the range we simulated are fairly common across a range of taxa. At the extreme low end of empirical neighborhood size estimates we see some flowering plants, large mammals, and colonial insects like ants. Species such as this have neighborhood size estimates small enough that spatial processes are likely to strongly influence inference.
These include some human populations such as the Gainj- and Kalam-speaking people of Papua New Guinea, in which the estimated neighborhood sizes in \citep{Rousset1997} range from 40 to 410 depending on the method of estimation. \revpoint{3}{7}
Many more species occur in a middle range of neighborhood sizes between 100 and 1000 -- a range in which spatial processes play a minor role in our analyses under random spatial sampling but are important when sampling of individuals in space is clustered. Surprisingly, even some flying insects with huge census population sizes fall in this group, including fruit flies (\textit{D.~melanogaster}) and mosquitoes (\textit{A. aegypti}). Last, many species likely have neighborhood sizes much larger than we simulated, including modern humans in northeastern Europe \citep{Ringbauer2017}. For these species demographic inference and summary statistics are likely to reflect minimal bias from spatial effects as long as dispersal is truly continuous across the landscape. While that is so we caution that association studies in which the effects of population structure are confounded with spatial variation in the environment are still sensitive to dispersal even at these large neighborhood sizes.
